---
title: "Using plotting and preparation functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using plotting and preparation functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows how to use the plotting-preparation and plotting functions in _vimcheck_.

Note that all data used here are placeholders.

```{r setup}
library(vimcheck)
```

## Compare demography

Users can check demographic alignment of burden data using `check_demography_alignment()`, then prepare it for plotting using `prep_plot_demography()`, and plot it using `plot_compare_demography()`.

```{r comparing_demography}
burden <- eg_burden_template

burden <- check_demography_alignment(burden, eg_wpp)
burden <- prep_plot_demography(burden)

plot_compare_demography(burden, 1)
```

## Examine age patterns

Users can check age patterns in burden data using `prep_plot_age()` and plotting using `plot_age_patterns()`.

Note that values are placeholders and you should expect to see real age-wise burden patterns look very different.

```{r age_patterns}
burden <- eg_burden_template
burden <- prep_plot_age(burden)

# manually set values as template default is NA, prevents ggplot warnings
burden$value_millions <- 1.0

plot_age_patterns(burden, 1)
```

## Global burden by decade

Users can check the global burden in each decade for each scenario using `prep_plot_burden_decades()` and `plot_global_burden_decades()`.

```{r global_burden_decades}
burden <- eg_burden_template
year_max <- 2100
burden <- prep_plot_burden_decades(burden, year_max)

# manually set values as template default is NA, prevents ggplot warnings
burden$value_millions <- 1.0

plot_global_burden_decades(burden, 1)
```

## Global burden timeseries

Users can check a timeseries of global burdens by scenario and age group.
In contrast with the plotting scheme above, `prep_plot_global_burden()` converts the burden data to long-format and transforms the data tibble into a [nested-tibble](https://tidyr.tidyverse.org/articles/nest.html).
This gives a tibble with as many rows as burden outcomes: cases, deaths, DALYs and YLLs, and a tibble giving the annual values by age for each burden outcome.

The function `plot_global_burden()` is intended to be applied row-wise, taking the burden outcome name (e.g. "cases") and the burden outcome data to plot a timeseries with values by age.

```{r global_burden_ts}
burden <- eg_burden_template
burden <- prep_plot_global_burden(burden)

# NOTE: expected use case is to loop over nested column DFs
# set values to a dummy placeholder
burden$burden_data[[1]]$value_millions <- 1

plot_global_burden(
  burden$burden_data[[1]],
  burden$burden_outcome[[1]],
  1
)
```
