---
title: "Pressure testing functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pressure testing functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows how to use functions added from the pressure testing report.

```{r setup}
library(vimcheck)
library(dplyr)
library(tidyr)
```

## Example data

Example impact data are taken from [_vimpact_](github.com/vimc/vimpact) and included in the package as [eg_impact].
This dataset holds projections for four countries, four diseases, and three modelling groups; combinations are shown below.

```{r impact_data}
eg_impact

# check countries
unique(eg_impact$country_name)

# check diseases
unique(eg_impact$disease)

# check modelling group
unique(eg_impact$modelling_group)

# check combinations
distinct(eg_impact, country_name, disease, modelling_group)
```

Data on WHO regions is provided as [who_subregions] to enable comparing countries with their regions.

```{r who_regions}
who_subregions
```

## Filtering impact data

Impact data can be filtered, or flagged for filtering, in different ways.

Flag data for filtering on touchstone using [filter_recent_ts()].

```{r filter_ts}
# assign dummy touchstones and scenario type for demo
df <- eg_impact
df$touchstone <- rep(
  c("202412", "202012"),
  each = nrow(df) / 2
)

test_scenario_types <- rep(
  c("default", "dummy"),
  each = nrow(df) / 2
)

# use a package default touchstone
DEF_TOUCHSTONE_NEW

# only touchstones from 2020 remain
filter_recent_ts(df, DEF_TOUCHSTONE_NEW)
```
